COFFEE: an Optimizing Compiler for Finite Element Local Assembly

### 2014-06-23 TACO-paper/coffee.tex
Still in the process of writing licm-licmap results

\begin{table*}[t]
\tbl{Performance improvement due to generalized loop-invariant code motion (\emph{licm} columns) on the Helmholtz, Diffusion and Burgers problems, varying the element shape (triangle, tetrahedron, prism) and the polynomial order ($p \in [1, 4]$), using the Lagrange element family, over the original, non-optimized code. The \emph{licm-ap} columns illustrate the effect of combining \emph{licm} with data alignment and padding. Results are shown for both the Sandy Bridge and the Xeon Phi architectures.}{
\scriptsize
\massmatrixtwodresultsnorms
}
\label{table:perf-results-licm}
\end{table*}

Table~\ref{table:perf-results-licm} illustrates the performance improvement obtained on the Sandy Bridge and Xeon Phi machines when \emph{licm}, data alignment, and padding are used. We distinguish between \emph{licm} and \emph{licm-ap}; the latter makes use of padding and data alignment. Inspection of assembly code generated by \texttt{icc} confirmed all limitations described in Section~\ref{sec:code-transf-licm}: only sub-expressions invariant to the inner loop are hoisted and their execution is not vectorized. Padding and data alignment enhance, in general, the quality of auto-vectorization. Sometimes the run-time of \emph{licm-ap} is similar to that of \emph{licm} because the element matrix size is already a multiple of the vector length, so no scalar reminder loop has to be introduced. Occasionally \emph{licm-ap} is slower than \emph{licm} (e.g. in Burgers $p=3$ on the Sandy Bridge). One possible explanation is that the large number of aligned temporaries introduced by \emph{licm} induce cache associativity conflicts.

>>>>

\begin{table*}[t]
\tbl{Performance improvement due to generalized loop-invariant code motion (\emph{licm} columns) on the Helmholtz, Diffusion and Burgers problems, varying the element shape (triangle, tetrahedron, prism) and the polynomial order ($p \in [1, 4]$), using the Lagrange element family, over the original, non-optimized code. The \emph{licm-ap} columns illustrate the effect of combining \emph{licm} with data alignment and padding. Results are shown for both the Sandy Bridge and the Xeon Phi architectures.}{
\scriptsize
\massmatrixtwodresultsnorms
}
\label{table:perf-results-licm}
\end{table*}
Table~\ref{table:perf-results-licm} illustrates the performance improvement obtained when \emph{licm} is used. The gain grows with the computational cost of the kernels: bigger loop nests usually benefit from the reduction in redundant computation, even though extra memory for the temporary arrays is required. Inspection of assembly code generated by \texttt{icc} confirmed all limitations described in Section~\ref{sec:code-transf-licm}: in the original version of the code, only sub-expressions invariant to the inner loop were hoisted, and their execution was not vectorized, which motivates the generally good speed ups achieved by \emph{licm}.

\subsection{Impact of Padding and Data Alignment}
Table~\ref{table:perf-results-licmap} shows the cumulative impact of \emph{licm}, data alignment, and padding over the original code. In the following, this version of the code is referred to as \emph{licm-ap}. Padding and data alignment enhance, in general, the quality of auto-vectorization. Occasionally the run-time of \emph{licm-ap} is close to that of \emph{licm}, since the element matrix size is already a multiple of the vector length, and no penalty for a scalar reminder loop has to be paid. Rarely \emph{licm-ap} is slower than \emph{licm} (e.g. in Burgers $p=3$ on the Sandy Bridge). One possible explanation is that the number of aligned temporaries introduced by \emph{licm} is so large to induce cache associativity conflicts.
